Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-23 paper/icml_drau/advup.tex

To obtain a timestep-invariant RL algorithm, a timestep-invariant
exploration scheme is required. 
For \emph{continuous} actions, \cite{ddpg} already introduced such a scheme, 
by adding an
\emph{Ornstein--Uhlenbeck} \cite{orn-uhl} (OU) process to actions. Formally, it is defined as
\begin{equation}
	\pi^\text{explore}(s_{k\deltat}, z_{k\deltat}) \deq
	\pi(s_{k\deltat}) + z_{k\deltat}
\end{equation}
with $z_{k\deltat}$ the discretization of a continuous-time OU process,
\begin{equation}
	dz_t = - z_t \,\kappa\, dt + \sigma \,\dbt.
	\label{eq:orn_uhl}
\end{equation}
where $B_t$ is a brownian motion, $\kappa$ a stiffness parameter and
$\sigma$ a noise scaling parameter. The discretized trajectories converge
to non trivial continuous time trajectories, exhibiting a Brownian
behavior with a recall force towards $0$.

This exploration can be extended to schemes of the form
\begin{equation}
  \label{eq:explore}
	a^k_\deltat = \pi^\text{explore}_\deltat(s^k_\deltat, z^k_\deltat)
\end{equation}
with $(z^k_\deltat)_{k\geq 0}$ a sequence of random variables independent from the $a$'s and $s$'s.
A sufficient condition for this policy to admit a continuous-time
limit is for the sequence
$z_\deltat$ to converge in law to a
well-defined continuous stochastic process $z$ as $\deltat$ goes to $0$.
% For instance, $\varepsilon$-greedy exploration naturally falls in this framework, \TODO{[[CT would like to give these details, LB thinks it is not necessary:]] with each $z^k_\deltat$
% being a pair composed of a bernoulli random variables $b^k_\deltat$ with probability of being $1$ equal to $\varepsilon$,
% and of a uniform categorical random variable $\xi^k_\deltat$ on actions. Exploratory actions are then selected as
% \begin{equation}
% 	a^k_\deltat = (1 - b^k_\deltat) \text{argmax}_{a'} Q^\pi_\deltat(s, a') + b^k_\deltat \xi^k_\deltat.
% \end{equation}}

Thus, for discrete actions we can obtain a temporally consistent
exploration scheme by taking
$z_\deltat$ to be a discretization of an $|\mathcal{A}|$-dimensional
continuous OU process, and setting
\begin{equation}
  \pi^\text{explore}(s^k_\deltat, z^k_\deltat)\deq \text{argmax}_{a'} \left(A_\psi(s^k_\deltat, a') + z^k_\deltat[a']\right)
\end{equation}
where $z^k_\deltat[a']$ denotes the $a'$-th component of $z^k_\deltat$. Namely,
we perturb advantage values by a random process before selecting an action. The
resulting scheme converges in continuous time to a nontrivial exploration
scheme.

>>>>

To obtain a timestep-invariant RL algorithm, a timestep-invariant
exploration scheme is required. 
For \emph{continuous} actions, \cite{ddpg} already introduced such a scheme, 
by adding an
\emph{Ornstein--Uhlenbeck} \cite{orn-uhl} (OU) random process to the
actions. Formally, this is defined as
\begin{equation}
	\pi^\text{explore}(s_{k\deltat}, z_{k\deltat}) \deq
	\pi(s_{k\deltat}) + z_{k\deltat}
\end{equation}
with $z_{k\deltat}$ the discretization of a continuous-time OU process,
\begin{equation}
	dz_t = - z_t \,\kappa\, dt + \sigma \,\dbt.
	\label{eq:orn_uhl}
\end{equation}
where $B_t$ is a brownian motion, $\kappa$ a stiffness parameter and
$\sigma$ a noise scaling parameter. The discretized trajectories of $z$ converge
to nontrivial continuous-time trajectories, exhibiting Brownian
behavior with a recall force towards $0$.

This exploration can be extended to schemes of the form
\begin{equation}
  \label{eq:explore}
	a_{k\deltat} = \pi^\text{explore}_\deltat(s_{k\deltat},
	z_{k\deltat})
\end{equation}
with $(z_{k\deltat})_{k\geq 0}$ a sequence of random variables independent from the $a$'s and $s$'s.
A sufficient condition for this policy to admit a continuous-time
limit is for the sequence
$z_{k\deltat}$ to converge in law to a
well-defined continuous stochastic process $z_t$ as $\deltat$ goes to $0$.
% For instance, $\varepsilon$-greedy exploration naturally falls in this framework, \TODO{[[CT would like to give these details, LB thinks it is not necessary:]] with each $z^k_\deltat$
% being a pair composed of a bernoulli random variables $b^k_\deltat$ with probability of being $1$ equal to $\varepsilon$,
% and of a uniform categorical random variable $\xi^k_\deltat$ on actions. Exploratory actions are then selected as
% \begin{equation}
% 	a^k_\deltat = (1 - b^k_\deltat) \text{argmax}_{a'} Q^\pi_\deltat(s, a') + b^k_\deltat \xi^k_\deltat.
% \end{equation}}

Thus, for \emph{discrete} actions we can obtain a consistent
exploration scheme by taking
$z_\deltat$ to be a discretization of an $(\#\mathcal{A})$-dimensional
continuous OU process, and setting
\begin{equation}
  \pi^\text{explore}(s_{k\deltat}, z_{k\deltat})\deq \text{argmax}_{a}
  \left(A_\psi(s_{k\deltat}, a) + z_{k\deltat}[a]\right)
\end{equation}
where $z_{k\deltat}[a]$ denotes the $a$-th component of $z_{k\deltat}$. Namely,
we perturb the \emph{advantage values} by a random process before selecting an action. The
resulting scheme converges in continuous time to a nontrivial exploration
scheme.
